---
title: "First Steps in `R`"
output: html_notebook
Name: Carlos A. Diaz
---

# Basic calculations

You can use R for basic computations you would perform in a calculator

```{r}
# Addition
56223+8953
# Division
4658/89
# Exponentiation
23^6
# Square root
sqrt(121)
# Logarithms
log(56)
```

Often you will want to test whether something is less than, greater than or equal to something.

```{r}
19 == 150
456 != 78946
2 <= 789
```

The _logical operators_ are `&` for logical **AND**, `|` for logical **OR**, and `!` for **NOT**. These are some examples:

```{r}
# Logical Disjunction (or) 
#FALSE | FALSE
#example of comparing two lists
c(0, 0) | c(0, 0)

# Logical Conjunction (and)
TRUE & FALSE

#example
x <- c(TRUE,2+3i)
t <- c(FALSE,2+3i)
print(x&t)

# Negation
#! FALSE
2 != 2

# Combination of statements
2500 < 4000 | 500 == 9000
```

# Assigning Values to Variables


In R, you create a variable and assign it a value using `<-` as follows

```{r}
#assigning variable and doing complex operation
dog1<-sin(((sqrt(56))*9))/2.5
(dog1*4)/0.5

```


To see the variables that are currently defined, use `ls` (as in "list")

```{r}
#this fucntion is to see the lists that are assigned in this document. 

ls()

#will give as output the variable "dog1" defined in the last step. 
```

To delete a variable, use `rm ` (as in "remove")

```{r}
#this function will delete the variable "dog1" previously defined. 

rm(dog1)
```

Either `<-` or `=` can be used to assign a value to a variable, but I prefer `<-` because is less likely to be confused with the logical operator `==`

# Vectors

The basic type of object in R is a _vector_, which is an ordered list of values of the same type. You can create a vector using the `c()` function (as in "concatenate").

```{r}
bar <- c(2, 5, 10, 2, 1) 
bar

#creating vectors
vec1<-c(5,8,9,7,5,6,1,2,45,45,89,8,9,9,2,1,8,9,8)
vec1

```

```{r}
baz <- c(2, 2, 3, 3, 3)
baz

#another vector example

vec2<-c(1,9,8,6,5,4,5,8,9,5,6,74,3,2,5,6,5,4,8)
vec2

```


There are also some functions that will create vectors with regular patterns, like repeated elements. 

```{r}
# replicate function
rep(2, 5)

#this function replicates the number 15, 10 times. 

rep(15,10)

# consecutive numbers
1:5

#count consecutive numbers, from 100 to number 2
100:2

# sequence from 1 to 10 with a step of 2
seq(1, 10, by=2)

#sequence from 1 to 91 in tenths
seq(1, 100, by=10)

ls()
```

Many functions and operators like `+` or `-` will work on all elements of the vector.

```{r}
# add vectors
bar + baz

#adding previous vectors from steps above
vec1 + vec2

# compare vectors
bar == baz

#comparing previous set vectors

vec1==vec2

# find length of vector
length(bar)

#fiding the lengths of the vectors
length(vec1)
length(vec2)


# find minimum value in vector
min(bar)
#finding minimun values in the vectors

min(vec1)
min(vec2)




# find average value in vector
mean(bar)

#finding the average/mean in the vectors

mean(vec1)
mean(vec2)

```

You can access parts of a vector by using `[`. Recall what the value is of the vector `bar`.

```{r}
bar

#recalling the vector
vec2

# If you want to get the first element:
bar[1]

#getting the second element of the vector. 
vec2[2]

```


If you want to get the last element of `bar` without explicitly typing the number of elements of `bar`, make use of the `length` function, which calculates the length of a vector:

```{r}
bar[length(bar)]

#calculating the lenght of the vector and getting the last element

vec2[length(vec2)]



```


You can also extract multiple values from a vector. For instance to get the 2nd through 4th values use

```{r}
bar[c(2, 3, 4)]

#extracting multiple values of the vectors
vec1[c(1,5,10)]
vec2[c(4,9,10)]

```


Vectors can also be strings or logical values

```{r}
quxx <- c("a", "b", "cde", "fg")


#creating a vector with only strings

string1 <- c("train", "PC", "car", "truck", "coffee", "sugar", "pen", "pencil")

vec1
vec2

```


# Data Frames

In statistical applications, data is often stored as a data frame, which is like a spreadsheet, with _rows as observations_ and _columns as variables_.

To manually create a data frame, use the `data.frame()` function.

```{r}
data.frame(foo = c(1, 2, 3), 
           bar = c("a", "b", "c"), 
           baz = c(1.5, 2.5, 3)) 


#creatinf a data frame using function with previous vectors. 

data.frame(vec3 = c(5,8,9,7,5,6,1,2,45,45),
           vec4 = c(1,9,8,6,5,4,5,8,9,5),
           string1 = c("train", "PC", "car", "truck", "coffee", "sugar", "pen", "pencil", "screen", "cell"))

```

Most often you will be using data frames loaded from a file. For example, load the results of a class survey. The function `load` or `read.table` can be used for this. 

# How to Make a Random Sample

To randomly select a sample use the function `sample()`. The following code selects 5 numbers between 1 and 10 at random (without duplication)

```{r}
sample(1:10, size=5)


#creating a sample of 10 number on from 1 to 100 scale 

sample(2:100, size = 10)


```


- The first argument gives the vector of data to select elements from.
- The second argument (`size=`) gives the size of the sample to select.

Taking a simple random sample from a data frame is only slightly more complicated, having two steps:

1. Use `sample()` to select a sample of size `n` from a vector of the row numbers of the data frame. 
2. Use the index operator `[` to select those rows from the data frame.


Consider the following example with _fake data_. First, make up a data frame with two columns. (`LETTERS` is a character vector of length 26 with capital letters “A” to “Z”; `LETTERS` is automatically defined and pre-loaded in `R`)

```{r}
bar <- data.frame(var1 = LETTERS[1:10], var2 = 1:10)
# Check data frame
bar


#Matching data with letters. 
chunk17<- data.frame(Column1 = LETTERS[1:10], Column2 = 1:10)
chunk17




```


Suppose you want to select a random sample of size 5. First, define a variable `n` with the size of the sample, i.e. 5

```{r}
n <- 5

#defining a variable and getting a sample from it. 
n.2 = c(0,1,2,3)
n.2 <-2



```

Now, select a sample of size 5 from the vector with 1 to 10 (the number of rows in `bar`). Use the function `nrow()` to find the number of rows in `bar` instead of manually entering that number. 

Use `:` to create a vector with all the integers between 1 and the number of rows in `bar`.

```{r}

samplerows <- sample(1:nrow(bar), size=5) 
# print sample rows
samplerows

#taking a sample of 4 from vector bar

test1 <- sample(4:nrow(bar), size=5) 
# print sample rows
test1


```



The variable `samplerows` contains the rows of `bar` which make a random sample from all the rows in `bar`. Extract those rows from `bar` with

```{r}
# extract rows
barsample <- bar[samplerows, ]
# print sample
print(barsample)


#extracting columns from vectors
testsample1 <- chunk17[2,]
print(testsample1)


```



The code above creates a new _data frame_ called `barsample` with a random sample of rows from `bar`.

In a single line of code: 

```{r}
bar[sample(1:nrow(bar), n), ]



#creating a 
chunk17[sample(1:nrow(chunk17),n),]



```

# Using Tables

The `table()` command allows us to look at tables. Its simplest usage looks like `table(x)` where `x` is a _categorical variable_.

For example, a survey asks people if they smoke or not. The data is 

_Yes, No, No, Yes, Yes_

We can enter this into R with the `c()` command, and summarize with the `table()` command as follows

```{r}
x <- c("Yes","No","No","Yes","Yes") 
table(x)

#summarizing categorical information

categ1 <-c("cut", "uncut","cut", "uncut","cut", "uncut","cut", "uncut","cut", "uncut","cut", "uncut","cut")

table(categ1)


```

# Numeric measures of center and spread

Suppose, CEO yearly compensations are sampled and the following are found (in millions)

12    .4    5     2     50    8     3     1     4     0.25

```{r}
sals <- c(12, .4, 5, 2, 50, 8, 3, 1, 4, 0.25)
# the average
mean(sals) 
# the variance
var(sals)
# the standard deviation
sd(sals)
# the median
median(sals)
# Tukey's five number summary, usefull for boxplots
# five numbers: min, lower hinge, median, upper hinge, max
fivenum(sals)
# summary statistics
summary(sals)

####numeric measures of center and spread

Mes1 <- c(56,48,45,56,12,56,89,12,4,5,45,78,65,32,1,56,48,59,26,23,21,11)

# the average
mean(Mes1) 
# the variance
var(Mes1)
# the standard deviation
sd(Mes1)
# the median
median(Mes1)
# Tukey's five number summary, usefull for boxplots
# five numbers: min, lower hinge, median, upper hinge, max
fivenum(Mes1)
# summary statistics
summary(Mes1)






```


### How about the _mode_? 

In R we can write our own _functions_, and a first example of a function is shown below in order to compute _the mode_ of a vector of observations `x`

```{r}
# Function to find the mode, i.e. most frequent value
getMode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
 }
```

As an example, we can use the function defined above to find the most frequent value in te vector `baz`

```{r}
# Most frequent value in baz
getMode(baz)


#finding the mode of vector Mes1

getMode(Mes1)
```



```{r}
```




