---
title: "Handout_Last"
output: html_notebook
Handout 15
Student: Carlos Diaz
---

Good afternoon all! First let us find out what our working directory is.

```{r}
getwd()
```

Today we are going to query the table quiz2 by using Transact-SQL inside the R chunks. First, we are going to need to install the package **sqldf**.

```{r}
install.packages("sqldf")

#########################################################################################

#getting new package for R to use SQL within R
########################################################################################


```


```{r}
library(sqldf)

####################################################################################

#Calling R library

###################################################################################

```


```{r}
data1<-read.csv("quiz2.csv",header = TRUE,sep = ",")


####################################################################################

#Reading file quiz2 from working directoy 

###################################################################################

```


```{r}
summary(data1)


####################################################################################

#Getting a summary of the data

###################################################################################
```

Let us run a query that returns every single record from the table data1.

```{r}
sqldf("SELECT * FROM data1")


####################################################################################

#Running sql in R. Select all records from data1

###################################################################################

```

Now I would like you to create a query that returns every single record (row) for the fields region and population.

#############################################################

#see answer below. 

##############################################################

```{r}

sqldf("SELECT region, pop FROM data1")

```





Let us now create a query that returns the first 10 records of the data1 table.

```{r}
sqldf('SELECT * FROM data1 LIMIT 10')

#####################################################

#create a query of all records from file data1 but limit to 10 first rows

##############################################


```


```{r}
sqldf("SELECT * FROM data1 ORDER BY infections DESC  LIMIT 5")

###############################################################################

#running a query of all data from file data1 sort by infections descending order and limit to the first 5 rows. 

##############################################################################

```

Interpret the output of the query ran above and run a similar query (limit 7) including the attributes **ipadddr** , **ufo2010** ,and **infections**.

################################################################################

#see answer below. 

###############################################################################


```{r}
sqldf("SELECT ipaddr, ufo2010, infections FROM data1 LIMIT 7")
```


```{r}
sqldf('SELECT region,infections FROM data1 WHERE infections > 1000 order by 2 desc')

################################################################################

#Select columns region and infections from data1 file which infections are greater than 1000 sort the result by 2 in descending order. 

###############################################################################


```

Now I would like you to inspect (create a query) income,region and population when the condition is income>=50,000. Make comments.

```{r}
sqldf('SELECT region,income, pop FROM data1 WHERE income >=50000 ORDER BY income')

#####################################################################################
#Here I am selecting all records from columns income, region and population from the file data1 and specifying the output when the income is larger than 50000. I added to order by the income. SQL will order ascending if not specified. 
#####################################################################################
```



```{r}
sqldf('SELECT * FROM data1 WHERE (income > 50000 AND infections > 2000) OR pop > 100000')
#####################################################################################
#Here the author is selecting all records from file data1 and specifying that income must be greater than 50000 and infections above 2000 if any of those conditions do not happened, then display population above 100000. 
#####################################################################################

```

Interpret the query ran above and create a query that recreates a similar case scenario (one that is critical for the success of this mini-project).

```{r}
#####################################################################################
#maybe the query could be the following
#sqldf('select infections from test_set where (infections =Yes) )
#####################################################################################
```



Let us run a query that returns every region and population with a number of infections between 1000 and 7000. 

```{r}
sqldf('SELECT region,pop FROM data1 WHERE infections between  1000 and 7000')
```

Interpret the query below and list a hypothetical case scenario in which this query would be useful.

```{r}
sqldf('SELECT * FROM data1 WHERE subregion LIKE "A%"')
##############################################################################
# this query reports all data from the data1 file. It specifies from the column subregion the records that start with capital A and it does not matter the rest. 

#this type of queries with commodin, are useful when we do not remember the complete name of the record we are looking for. 


#############################################################################
```

Last,let us find the average number of infections in our table.

```{r}
sqldf("SELECT AVG(infections) FROM data1")

################################################################################
#in this case we are reporting the average infections from file data1. 
################################################################################

```

Let us inspect the new table defined below data2.

```{r}
data2<-read.csv("C:\\Users\\carlo\\Downloads\\finalexam.csv",header = TRUE,sep = ",")
```

Let us apply summary to data2 and examine the output.
```{r}
summary(data2)
```

Next, I would like you to complete the following steps:

1-Run a query that returns full details for every record from the west region.

```{r}
sqldf('select * FROM data2 WHERE(region="West")')
```


2-Run a query that returns full details for every record from the east region where the number of infections is greater than 1500.

```{r}
sqldf('select * from data2 where(region="East" and infections>1500)')
```

3-Run a few queries that can potentially help us to make the point that:
   a.Region affects the number of infections
```{r}
sqldf('select region, infections from data2 where(infections >=1000)')

######################################################################################
#Here I am trying to report that region affects the number infections. 
#####################################################################################

```

   
   b.Population affects the number of infections
```{r}
sqldf('select pop, infections from data2 where(infections >=1000)')

######################################################################################
#Here I am trying to report that population affects the number infections. 
#####################################################################################
```

   c.Income affects the number of infections
```{r}
sqldf('select income, infections from data2 where(infections >=1000)')

######################################################################################
#Here I am trying to report that income affects the number infections. 
#####################################################################################
```

   d.ufo2010 affects the number of infections
```{r}
sqldf('select ufo2010, infections from data2 where(infections >=1000)')

######################################################################################
#Here I am trying to report that ufo2010 affects the number infections. 
#####################################################################################
```

   e. The different interactions of region and income, and income and population affects the number of infections.
```{r}
sqldf('SELECT region, income, infections
         FROM data2
                WHERE region = "East"
                
      Union
      SELECT income, pop, infections
                FROM data1 
                WHERE (infections>1000)')

##############################################################################################
#in this code I am joining to variables from data1 and data2 in order to make 2 interactions from 2 different files and demonstrate that region and income affect infections. 
################################################################################################
```
   
   
Explain every single query you include in this part of the handout.



