---
title: "Handout_11_Reg"
output: html_notebook
Student: Carlos Diaz
---

---
title: "Handout_11_reg"
output: html_notebook
---




```{r}
getwd()
```



```{r}
# make sure the packages for this chapter
# are installed, install if necessary
pkg <- c("ggplot2", "scales", "maptools",
              "sp", "maps", "grid", "car" )
new.pkg <- pkg[!(pkg %in% installed.packages())]
if (length(new.pkg)) {
  install.packages(new.pkg)  
}

#############################################################################

# adding gg plot to this handout. 
#############################################################################
```



```{r}
# read the CSV with headers
regression1<-read.csv("incidents.csv", header=T,sep =",")

#############################################################################

# Reading the file incidents from downloads folder. 
#############################################################################

```



```{r}
View(regression1)


#############################################################################

# Viewing the file. 
#############################################################################

```



```{r}
summary(regression1)


#############################################################################

# Taking a look at the file to see the data.  
#############################################################################

```


```{r}
str(regression1)


#############################################################################

# Taking a look at the file to see the data.  
#############################################################################
```


```{r}
regression1$population <- as.numeric(gsub(",","",regression1$population))
regression1$population


#############################################################################

# Applying function to column population gsub (replaces all matches of a string, if the parameter is a string vector, returns a string vector of the same length and with the same attributes (after possible coercion to character) and saving it in a variable.  
#############################################################################



```

```{r}
str(regression1$population)
```


```{r}
regression2<-regression1[,-1]#new data frame with the deletion of column 1 

#############################################################################

# Getting a new data frame and deleting the first column. 
#############################################################################

```



```{r}
head(regression2)

#############################################################################

# Looking how the data looks after deleting the column. 
#############################################################################

```

```{r}
View(regression2)
```



```{r}
str(regression1)
```



```{r}
reg.fit1<-lm(regression1$incidents ~ regression1$population)

#############################################################################

# Applying the linear model regression on the column incidents and column population 
#############################################################################
```





```{r}
summary(reg.fit1)
#############################################################################

# Looking at the regression 

#############################################################################
```


```{r}
reg.fit2<-lm(incidents ~ zone+population, data = regression1)

#############################################################################

# Applying the linear model regression on the column incidents and columns population and zone and saving it on variable reg.fit2. 
#############################################################################

```



```{r}
summary(reg.fit2)
```



```{r}
regression1$zone <- ifelse(regression1$zone == "west", 1, 0)

#############################################################################

# In this code applying a conditional to column zone of regression file. To bring all values of West to a new varianble as 1 otherwise, bring the value as 0.  
#############################################################################
```{r}
View(regression1)
```



```{r}
str(regression1)
```


```{r}
#regression1$zone<-as.integer((regression1$zone),replace=TRUE) was not necessary

#############################################################################

# This is a comment and they did not run it. 
#############################################################################
```





```{r}
interaction<-regression1$zone*regression1$population
#############################################################################

# In this code they are getting an interaction value by multiplying zone by population.  
#############################################################################


```

```{r}
View(interaction)
```


```{r}
reg.fit3<-lm(regression1$incidents~interaction+regression1$population+regression1$zone)

#############################################################################

# In this code they are running a linear model in order to fit the data of incidents Vs interactions plus population and zone in the population regression. 
#####################################################################################


```



```{r}
summary(reg.fit3)


#############################################################################

# The model does not seem to have a very good P-values over 0.05 alpha and overall the Rsquared is over 50%, therefore, they decided to run the model one more time but dropping a variable. 
#############################################################################


```

```{r}
reg.fit4<-lm(regression1$incidents~interaction)

#############################################################################

#this model they are running the incidents Vs the interactions. 

###################################################################



```


```{r}
summary(reg.fit4)

#############################################################################

# In this model, the data seems to fit better in the model. The interaction seems to have a better P-value at 0.01, which is significant at 0.05 alpha and the Rsquared is on the 38%.  
#############################################################################

```


