---
title: "Handout_9"
output:
  html_document:
    df_print: paged
    
#Student: Carlos Diaz
---

##Chapter IV,Part I

Welcome to Lecture#7 (Handout#9). Let us start by finding out our working directory.

```{r}
#getwd()
setwd("C:/Users/carlo/Desktop/CIS3368")
```



Now, let us upload the data sets needed for this handout into our project, **reputation.data** and **dest.ips**.



```{r}
av <- read.csv("reputation.data",sep = "#",header =FALSE )
##############################################################################

#The first part is about getting the working directory for R
#second, changing the working directory to the CIS class
#third, reading the reputationdata.csv, and putting it in a variable (AV)


##############################################################################
```

Make comments about the chunk of code ran above.

```{r}
View(av)
################################################################

#here we view the file that we just read

###############################################################
```

Make comments about the chunk of code ran above.

```{r}
#Let us proceed to create a header for each column.

colnames(av) <- c("IP", "Reliability", "Risk", "Type",
                  "Country", "Locale", "Coords", "x")

################################################################

#Here we createa a header for the file we just read. 

#################################################################
```


Make comments about the chunk of code ran above.

```{r}
View(av)
#########################################################################

#Here we view the file with the new tittles. 

#######################################################################
```

Make comments about the chunk of code ran above.

```{r}
# create a vector of lat/long data by splitting on ","
av.coords.vec <- unlist(strsplit(as.character(av$Coords), ","))
#########################################

#Here we split the data into a new vector including the coordinates. 

########################################
```



```{r}
View(av.coords.vec)
############################################

#here we view the vector. 

##########################################
```


```{r}
# convert the vector in a 2-column matrix
av.coords.mat <- matrix(av.coords.vec, ncol=2, byrow=TRUE)

########################################################################

#here we split the colum into a 2 column matrix. 

#######################################################################

```


```{r}
View(av.coords.mat)

####################################


#here we view the new vector. 

######################################
```



```{r}
# project into a data frame
av.coords.df <- as.data.frame(av.coords.mat)
#####################################################


#here we add the data to a new dataframe. 


#####################################################
```



```{r}
View(av.coords.df)

##############################################

#here we view the new vector. 

##############################################
```

Make comments about the chunk of code ran above.


```{r}
# name the columns 
colnames(av.coords.df) <- c("lat","long")

###############################################

#naming the columns on the new vector that we just created. 

#############################################


```

Make comments about the chunk of code ran above.

```{r}
# convert the characters to numeric values
av.coords.df$long <- as.double(as.character(av.coords.df$long))
av.coords.df$lat <- as.double(as.character(av.coords.df$lat))
##############################################################

#here we convert the characters to numeric values and add to a respective variable. 

##############################################################

```

Make comments about the chunk of code ran above.


```{r}
# requires packages: ggplot2, maps, RColorBrewer, scales
# requires object: av.coords.df (4-3)
# R code to extract longitude/latitude pairs from AlienVault data
# need plotting and mapping functions plus colors
#install.packages("ggplot2")
#install.packages("maps")
#install.packages("RColorBrewer")
#install.packages("scales")
```




```{r}
library(ggplot2)
library(maps)
library(RColorBrewer)
library(scales)
#Which of the libraries listed above was previously used in this course?

######################################

#all of the previous libraries have been used. 


######################################
```




```{r}
# extract a color palette from the RColorBrewer package
set2 <- brewer.pal(8,"Set2")
```



```{r}
# extract the polygon information for the world map, minus Antarctica
world <- map_data('world')
world <- subset(world, region != "Antarctica")
#Do you see the utility of the code ran above?Please explain.

############################################3

#getting the map of the world
#getting the map excluding Antartica

############################################3

```



```{r}

# plot the map with the points marking lat/lon of the geocoded entries
# plotting ~200K takes a bit of time
gg <- ggplot()
gg <- gg + geom_polygon(data=world, aes(long, lat, group=group), 
                        fill="white")
gg <- gg + geom_point(data=av.coords.df, aes(x=long, y=lat),  
                      color=set2[2], size=1, alpha=0.1)
gg <- gg + labs(x="", y="")
gg <- gg + theme(panel.background=element_rect(fill=alpha(set2[3],0.2), 
                                              colour='white'))
gg
```


Make comments about the chunk of code ran above.In addition, interpret the output and discuss your findings.
##########################################################################################

#regarding the code, it is plotting the infections around the world using the coordinates previously separated from the main file. 

#fidings: It looks like East of the United States, West around the California area, Western Europe and SouthWest Asia seems to be the most affected area of infections. 




######################################################################################3



```{r}
# read in list of destination IP addresses siphoned from firewall logs
dest.ips <- read.csv("dest.ips", col.names= c("IP"))

##########################

#This code is putting on a variable the destination IP addresses found from the firewall
#######################################################################################
```
 


```{r}
#take a look at the reliability of the IP address entries
# (you could also plot a histogram)
table(av[av$IP %in% dest.ips$IP, ]$Reliability)

###########################################################################

#Plotting a frequency between the frequency and the IP/reliability 
hist (av[av$IP %in% dest.ips$IP, ]$Reliability)

#####################################################################

```

Discuss your findings. 

It seems that the most frequency lies between the 2 an 3. Followed by 6 and later 4. It also seems by looking at the map that there is no an area that it has not been affected by the problem. There sesm to be a low reliability of the system and high risk of spreading the infection further. 

