---
title: "Handout 7"
output: html_notebook
Student: Carlos Diaz

---
Please let us use this handout to discuss Associations /Correlations. I would like you to make comments as you run the code we have listed in this handout. In your submission file you should include commments that show a clear understanding of the correlation concept.

```{r}
mfrow <- par()$mfrow
mar <- par()$mar
oma <- par()$oma

#####################################################################

#here the function par() will prepare the data frames for graphical parameters. 

##########################################################


```



```{r}
par(mfrow=c(2, 3), mar=c(0, 0, 2, 0), oma=c(1, 1, 1, 1))
set.seed(2)
x <- runif(500, min=0, max=pi)

#####################################################################

#previous code is in order to delimit the data frame and run 500 random points.  

##########################################################
```




```{r}
y <- rnorm(500, mean=x, sd=0.2)
plot(x,y, pch=19, cex=.8, col="#666699CC", axes=FALSE, xlab="", ylab="", 
     main=paste("Designation: ", round(cor(x,y), 2)))


#####################################################################

#previous code is in order to delimit the data frame and run 500 random but following normality points. There is specific code for no labels. The code also runs a correlation (that is also being rounded) between variables "x" and "y' that is in this case there are labeling as "designation". 

##########################################################
```

#################################################################################


#here we find a positive correlation between the two variables. One of the variables increases as the other increases as well in tandem.

#############################################################################
```

Make comments about the output as well as the syntax related to the chunk ran above.

```{r}
y <- rnorm(500, mean=-x, sd=0.3)
plot(x,y, pch=19, cex=.8, col="#666699CC", axes=FALSE, xlab="", ylab="", 
     main=paste("Correlation: ", round(cor(x,y), 2)))
     
####################################################################################

#There is a negative or inverse correlation in this plot. One variable decreases as the other increases. This is an example of negative correlation. 

#####################################################################################
     
     
```


```{r}
y <- rnorm(500, mean=sin(x), sd=0.2)
plot(x,y, pch=19, cex=.8, col="#666699CC", axes=FALSE, xlab="", ylab="", 
     main=paste("Correlation: ", round(cor(x,y), 2)))

########################################################################################

#this is an example of non linear relationship between two variables or an example of curvelinear relationship. 

########################################################################################


```




```{r}
y <- rnorm(500, mean=x, sd=2)
plot(x,y, pch=19, cex=.8, col="#666699CC", axes=FALSE, xlab="", ylab="", 
     main=paste("Correlation: ", round(cor(x,y), 2)))
     
###################################################################################

#this might be an example of a very small positive correlation between two variables. Some outliers are observed but it does appear as an positive outward but small correlation.

#################################################################################
     
```



```{r}
y <- rnorm(500, mean=-x, sd=1)
plot(x,y, pch=19, cex=.8, col="#666699CC", axes=FALSE, xlab="", ylab="", 
     main=paste("Correlation: ", round(cor(x,y), 2)))
     
##################################################################################
 #this might be an example of a very small negative correlation between two variables. Some outliers are observed but it does appear as an negative downward but small correlation.
 
 
################################################################################   
     
```



```{r}
y <- runif(500, min=0, max=pi)
plot(x,y, pch=19, cex=.8, col="#666699CC", axes=FALSE, xlab="", ylab="", 
     main=paste("Correlation: ", round(cor(x,y), 2)))
     
################################################################################
# this scatterplot does not seem to have a linear trend (either positive or negative). This is possible a zero correlation or near zero correlation plot. 


################################################################################
     
```





