---
title: "Getting Started with R,Part II"
output:
  html_notebook: default
  pdf_document: default
  Student: Carlos Diaz
---

Let us continue getting started with R as we start discussing important statistical concepts.


## Case-scenario 1
Kate must take four quizzes in a math class. If her scores on the first three quizzes are 71, 69, and 79, what score does she need on the final quiz for her overall mean to be at least 70?

## Solution

Given that 
$x_1 = 71, x_2 = 69, x_3 = 79$

we want to find $x_4$ such that the mean (average) grade is 
$\bar{x} >= 70$

Notice that in this case $n = 4$.

According to the information above:
$70 \times 4 = 71 + 69 + 79 + x_4$

so when $x_4 = 61$, the quiz average will be 70.


```{r}
# Grades so far
grades_before <- c(71, 69, 79)
# Average quiz grade wanted
wanted_grade <- 70
# Number of quizzes
n_quizzes <- 4
# Needed grade on quiz 4
x_4 <- n_quizzes*wanted_grade - sum(grades_before)
x_4

#These are Frank Grades following scenario #1
grades_before2 <- c(41,69,63,94,99)
#average quiz wanted
wg2 <- 70
#Number of quizes
n_quizzes2 <-c(6)
#needed grade for the quiz 6 =70
X_42<- n_quizzes2*wg2 - sum(grades_before2)
X_42

Frank_grades <-c(41,69,63,94,99,54)



```

According to the calculations above, Kate must score 61 or better on the final quiz to get an average quiz grade of at least 70.


We could confirm this, by using the function `mean()` in `R`

```{r}
# Quiz grades
kate_grades <- c(71, 69, 79,61)
# Find mean
mean(kate_grades)
# Find standard deviation
sd(kate_grades)
# Find maximum grade
max(kate_grades)
# Find minimum grade
min(kate_grades)
```








We can also use the `summary()` function to find basic statistics, including the median!

```{r}
summary(kate_grades)
```


Next, I would like you to explain in detail every single task we completed above. In addition, let us deal with a similar case scenario and complete every single task we execute in Case-scenario 1.

######################################################################################################################
Step 1: Using mean/average formula

What is happening with Kate scores is a substitution in the average/mean formula. The formula to calculate the average/mean is average =(x1,x2,x3,x4)/n(number of total scores). Since we already know the values of the first 3 tests, we substitute the values and isolating x4. Therefore, the formula would be: (average(desired grade)*N)-addition of (x1,x2,x3) = x4. Therefore, by substituting the values, it would look like this:

                (70*4)-(71,69,79) = x4
                               61 = x4  

Step 2:Using function mean

In this case, the desired grade is used to verified the number. There is a list created with Kate's grades kate_grades <- c(71, 69, 79,61), then the standard deviation, the maximum and the minimum numbers are created using R functions and 61 is the final number.  

Step3: Using median
This step is basically the same as step 2. However, the function summary is used to get the same values as step 2. It is a faster process. 




#######################################################################################################################

Frank must take six quizzes in a Physics class. If his scores on the first five quizzes are 41, 69,63,94, and 99, what score does he need on the final quiz for his overall mean to be at least 70?

See answer above

#############

###Now let us go back to Case-scenario 1

Another useful function is `quantile` to find 
```{r}
# the 25% 
quantile(kate_grades, 1/4)
# the 75%
quantile(kate_grades, 3/4)
# the function IQR finds the interquantile range
# IQR(x) = quantile(x, 3/4) - quantile(x, 1/4)
IQR(kate_grades)



################################################
#these are X_4 Frank Scores
Frank_grades

 # the 25% 
quantile(Frank_grades, 1/4)

# the 75%

quantile(Frank_grades, 3/4)

# the function IQR finds the interquantile range
# IQR(x) = quantile(x, 3/4) - quantile(x, 1/4)
IQR(Frank_grades)






```


Make comments about the output and run a similar query using Frank_grades

####################################################################################################

Using Quantile: In this case, the R function used is Quantile. For this, all the grades from the student are needed. See exmaple above. 



#####################################################################################################


# Case-scenario 2


The average salary of 10 men is 72,000 and the average salary of 4 women is 84,000. Find the mean salary of all 14 people.

## Solution

We can easily find the joined mean by adding both mean and dividing by the total number of people.

Let $n_1 = 10$ denote the number of men, and $y_1 = 72000$ their mean salary. Let $n_2 = 4$ the number of women and $y_2 = 84000$ their mean salary.
Then the mean salary of all 16 individuals is:
$\frac{n_1 x_1 + n_2 x_2}{n_1 + n_2}$

We can compute this in R as follows:

```{r}
n_1 <- 10
n_2 <- 4
y_1 <- 72000
y_2 <- 84000
# Mean salary overall
salary_ave <-  (n_1*y_1 + n_2*y_2)/(n_1+n_2)
salary_ave

###############################################################################################

#this is my answer
#using the formula above.

n_11 <- 10
n_22 <- 10
y_11 <- 95000
y_22 <- 75000
# Mean salary overall
salary_ave2 <-  (n_11*y_11 + n_22*y_22)/(n_11+n_22)
salary_ave2


###########################################################################################

```

Solve a similar problem by changing number of men and women as well as the average income for each group. Make comments about the output.
#################################################################################################
Answer: What is happening in the first example, the salary of men is lower than women, although women's salary is higher than men's there is a disparity in the number of subjects. So, the data seems to be biased towards women in this case and not very significant.  

On my example, I made both number of women and men the same, so I could get a better average between both genders at $85,000. 


#################################################################################################


# Case-scenario 3

The frequency distribution below lists the results of a test given in Professor Wang's String theory class.

Score   |   Number of students
--------|----------------------
10      | 5
9       | 10
8       | 6
7       | 8
6       | 3
5       | 2



1. Find the mean,the median and the standard deviation of the scores.

2. What percentage of the data lies within one standard deviation of the mean?

3. What percentage of the data lies within two standard deviations of the mean?

4. What percent of the data lies within three standard deviations of the mean? 

5. Draw a histogram to illustrate the data. 

## Solution

The `allScores.csv` file contains all the students' scores in the quiz. We can read this file in `R` using the `read.csv()` function (hint:First create a csv file with 6 rows and 2 columns)

```{r}
scores <- read.table("allScores.csv", header = TRUE, sep = ",")
WangScores <- scores$Score
############################################################################

#Comments in this code we are looking to read a file .csv with headers, separated by commnas. 


##################################################



#Make comments about the code we just ran above.  


#1. To find the mean and the standard deviation

# Mean 
Scores_mean  <- mean(WangScores)
Scores_mean
# Median
Scores_median <- median(WangScores)
Scores_median
# Find number of observations
Scores_n <- length(WangScores)
# Find standard deviation
Scores_sd <- sd(WangScores)


#######################################################################################

#frank Grades

Frank_grades



# Mean 
Scores_mean2  <- mean(Frank_grades)
Scores_mean2
# Median
Scores_median2 <- median(Frank_grades)
Scores_median2
# Find number of observations
Scores_n2 <- length(Frank_grades)
# Find standard deviation
Scores_sd2 <- sd(Frank_grades)



#####################################################################################







```

2. What percentage of the data lies within one standard deviation of the mean?

```{r}
scores_w1sd <- sum((WangScores - Scores_mean)/Scores_sd < 1)/ Scores_n
# Percentage of observation within one standard deviation of the mean
scores_w1sd
## Difference from empirical 
scores_w1sd - 0.68

#################################################################

#Frank_grades

scores_w1sd2 <- sum((Frank_grades - Scores_mean2)/Scores_sd2 < 1)/ Scores_n2
# Percentage of observation within one standard deviation of the mean
scores_w1sd2
## Difference from empirical 
scores_w1sd2 - 0.68



#####################################

```

3. What percentage of the data lies within two standard deviations of the mean?

```{r}
## Within 2 sd
scores_w2sd <- sum((WangScores - Scores_mean)/ Scores_sd < 2)/Scores_n
scores_w2sd
## Difference from empirical 
scores_w2sd - 0.95
####################################################################
#Frank_grades

## Within 2 sd
scores_w2sd2 <- sum((Frank_grades - Scores_mean2)/ Scores_sd2 < 2)/Scores_n2
scores_w2sd2
## Difference from empirical 
scores_w2sd2 - 0.95






####################################################

```


4. What percent of the data lies within three standard deviations of the mean?

```{r}
## Within 3 sd 
scores_w3sd <- sum((WangScores - Scores_mean)/ Scores_sd < 3)/Scores_n
scores_w3sd
## Difference from empirical 
scores_w3sd - 0.9973

###############################################

#Frank_grades

## Within 3 sd 
scores_w3sd2 <- sum((Frank_grades - Scores_mean2)/ Scores_sd2 < 3)/Scores_n2
scores_w3sd2
## Difference from empirical 
scores_w3sd2 - 0.9973





##############################################




```

Explain the implications of the results obtained in this problem. In addition, create a similar query but this time addressing Frank_Scores


5. Draw a histogram

```{r}
# Create histogram
hist(WangScores)


#######################################################

#Frank_grades create a histogram

hist(Frank_grades)


####################################




```

Explain the output and create a similar histogram for Frank_Scores.








